---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ducksf

<!-- badges: start -->
<!-- badges: end -->

`ducksf` extends the `sf` package to work with `duckdb` databases.

So far only `st_read()` is implemented partially, but future work will include using the `st_write()`. By writing methods for the `st_write` and `st_read` generics, hopefully much functionality of `sf` will be unlocked (such as using `sf`'s own `read_sf()` generic as below.)

Database-side evaluation for other `st_*` functions and `dbplyr` SQL generators for spaital queries is also an ambition of the project.

## Example

This is a basic example which shows reading an simple feature collection from duckdb. So far this is the only function implemented, and currently does not identify the geometry column(s) automatically.

```{r example, message = FALSE, warning = FALSE}
library(ducksf)

library(DBI)
library(duckdb)
library(sf)
```

```{r, message = FALSE, warning = FALSE}
# create database and create table with geometry
conn <- dbConnect(duckdb())

dbSendQuery(conn, "install spatial")
dbSendQuery(conn, "load spatial")
dbSendQuery(conn, "create table x (id int, geom geometry)")
dbSendQuery(conn, "insert into x values (1, st_point(1,1))")
```

```{r}
# read the table into R
read_sf(dsn = conn, query = "select * from x")
```

Note `duckdb`'s spatial extension currently does not supoort encoding the CRS into the geometry. This may come in the future, but in the meantime we may wish to explore ways of storing a metadata table in the duckdb database when writing `sf` objects. 

